|

### 2025-05-04 doc/paper.qmd
Template update - paper.qmd

---
title: |
  | How Much New Information Is There in Earnings?
  | Corporate Decision-Making and Quantitative Analysis
  | Winter 2024/25 - Individual Report\vspace{1cm}
author:
  - name: Mel Mzv
    email: m@student.hu-berlin.de
    affiliations:
      - Humboldt-Universität zu Berlin  
date: today
date-format: MMM D, YYYY [\vspace{0cm}] # Adjusted the vertical space after the date from 1cm
abstract: |
  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects. Leveraging data from the CRSP and Compustat databases, alongside Worldscope and Datastream via WRDS, this repository showcases a reproducible workflow integrating Python scripts for data preparation, analysis, and visualization. Integrating multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively. The project replicates and extends Ball and Shivakumar (2008) to analyze the informativeness of quarterly earnings announcements and their contribution to share price movements, highlighting their critical role in investment decisions and impact on investors and analysts. Key tasks include replicating and comparing original results, extending the analysis to 2007–2023, and applying the methodology to a non-U.S. country. The project also documents research design choices, discusses variations between original and reproduced results, and provides insights into earnings informativeness across different timeframes and jurisdictions. Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings. This code base, adapted from TREAT, demonstrates how the template applies to this project and serves as a structured guide for reproducible empirical research in accounting.
  <!--
  | \vspace{6cm} commenting away to decrease distance
  -->

bibliography: references.bib
biblio-style: apsr
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: true
    lof: true  # Enables List of Figures
    lot: false  # Disables List of Tables
fig_caption: yes
fontsize: 11pt
ident: yes
always_allow_html: yes
number-sections: true 
header-includes:
  - \usepackage[nolists]{endfloat}    
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=24pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
  - \usepackage[margin=1in]{geometry}
---

@@

# List of Abbreviations
\setlength{\parindent}{0em}
**BHR**: Buy-and-Hold Returns   
**CRSP**: Center for Research in Security Prices  
**CQA**: Corporate Decision-Making and Quantitative Analysis  
**DS**: Datastream  
**DV**: Dependent Variable  
**HU**: Humboldt-Universität zu Berlin  
**IDE**: Integrated Development Environment  
**IV**: Independent Variable  
**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
**US**: United States  
**WRDS**: Wharton Research Data Services  
**WSCP**: Worldscope

@@

# Introduction {#sec-introduction}
The aim of this paper is to illustrate the use of open science tools in empirical accounting research. This project builds on the methodological framework established in the Corporate Decision-Making and Quantitative Analysis (CQA) course, which explored a range of empirical research methods—including archival analysis, field experiments, and survey-based approaches—to develop a comprehensive understanding of corporate decision-making and quantitative analysis. Expanding on prior empirical research conducted in Assignments I and II, this paper extends the empirical focus of the course. Assignment I examined audit market concentration in the EU using Transparency Reports data, offering insights into the dominance of major audit firms and the structure of the European audit market. Assignment II explored the economic implications of corporate financial reporting, focusing on @graham_economic_2005 to analyze managerial decision-making, earnings management strategies, and voluntary disclosure practices.

Building on these foundations, Assignment III investigates the informativeness of quarterly earnings announcements and their contribution to annual share price movements. The project replicates and extends @ball_how_2008, assessing the extent to which earnings releases serve as a source of new information for financial markets. By integrating multiple datasets from WRDS and applying an event-study methodology, this study aims to evaluate the role of earnings disclosures in market efficiency, showcasing the valuation purpose of accounting. At the same time, earnings reports may play a crucial role in contracting by influencing managerial compensation, debt agreements, and other contractual mechanisms tied to financial performance [@ball_how_2008]. The integration of multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively.

Beyond empirical replication, the project also documents research design choices, replication steps, and explicit assumptions made whenever the original paper was unclear on how to proceed. It discusses variations between original and reproduced results and provides insights into the timeliness and market impact of earnings announcements across different timeframes and jurisdictions. Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings through an alternative methodological lens. This project leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to establish an open science-oriented infrastructure, ensuring transparency and replicability for empirical accounting research.

Earnings announcements provide investors with valuable information about a firm’s market value, with stock prices reacting significantly when earnings news deviate from expectations [@fink_review_2021, p.2]. The seminal study by @ball_empirical_1968 was the first to document this relationship, showing that stock prices anticipate earnings surprises, with most of the market reaction occurring before the official announcement, suggesting that earnings reports primarily confirm rather than introduce new information. Over the past decades, more than a thousand studies have examined the interplay between capital markets and financial statements, a research stream that originated with Ball and Brown (1968)  [@kothari_capital_2001].

Building on this foundation, the original study by @ball_how_2008 investigates the extent to which quarterly earnings announcements contribute new information to the market, assessing their role in shaping annual share price movements. By estimating the R² from regressions of annual stock returns on earnings announcement window returns, the paper quantifies the informativeness of earnings releases, finding that they account for only 1% to 2% of total annual volatility. This challenges the assumption that earnings provide substantial new information and instead suggests a confirmatory role in financial reporting. These findings align with the semi-strong form of market efficiency, which suggests that stock prices already reflect all publicly available information, including prior signals about earnings. Unlike traditional event studies that assume market efficiency, @ball_how_2008 allow for potential market mispricing by not imposing a unit slope restriction in their regressions, meaning that their approach can capture under- or overreaction to earnings announcements. They also emphasize that extending the event window beyond the earnings announcement period would shift the research focus from earnings “surprise” to broader financial reporting effects, such as its role in debt contracting and executive compensation.

While @ball_how_2008 provide key insights into the informativeness of earnings announcements, their study was conducted within a specific timeframe and market context. Since then, financial markets have undergone significant changes due to regulatory reforms, economic crises, and technological advancements in financial information processing. Moreover, the extent to which earnings informativeness varies across international markets remains underexplored. By replicating their study with updated data from 2007–2023 and applying the methodology to Canada as a non-U.S. country, this project reassesses the robustness of their findings. Through this approach, it critically evaluates the generalizability of the original results.

The paper is structured into sections corresponding to Tasks 1–3 (@sec-task1, @sec-task2, @sec-task3), each detailing the research design choices and assumptions, documenting the replication steps and results. @sec-task1 follows an exact replication approach by replicating key tables and figures from the original study using US data (1972–2006) to establish baseline results. @sec-task2 extends the analysis through empirical generalization, applying the same methodology to 2007–2023 US data to compare how results evolve over time. @sec-task3 employs a generalization and extension approach by applying the methodology to a non-US country, requiring adjustments for different databases and market structures. This framework aligns with the replication taxonomy outlined by @salterio_replication_2022. Finally, @sec-task4 sketches the survey design as a non-archival study that allows evaluating the key findings of the seminal paper, following Bloomfield’s (2016) triangulation principle. The concluding remarks are provided in @sec-conclusion.

# Task 1 - Replication of Key Tables and Figures {#sec-task1}
Task 1 is the most comprehensive, replicating the methodology of key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results, ensuring alignment with the original findings before extending the analysis in @sec-task2. By reconstructing Table 1 (Panel A), Table 2 (all panels), and Figure 1 (all panels), the task evaluates the robustness of the original results and identifies potential discrepancies, offering insights into data consistency, methodology, and market dynamics over the examined period. The use of daily CRSP data is essential for capturing short-term stock price movements around earnings announcements, ensuring precise event window calculations. The discussion of replication findings is presented in @sec-results1.

## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
Following @ball_how_2008, I focus on earnings announcements and stock returns from 1972 to 2006, ensuring alignment with the original study’s methodology. This replication mirrors the research design, maintaining consistency in sample selection, event window definitions, and return calculations.

Based on the original paper, I do not gather analyst expectations or estimate earnings deviations, requiring only stock return data and earnings announcement dates, simplifying data collection.

To ensure clarity where @ball_how_2008 lacks explicit guidance, I impose the following assumptions:

1.	Earnings announcement windows are defined as the three-day period surrounding the release (-1 to +1). According to @center_for_research_in_security_prices_crsp_crsp_2024 guide,since CRSP excludes weekends and holidays, if Day 0 falls on a non-trading day, it shifts to the next trading day. Similarly, Day -1 or +1 adjusts to the nearest available trading day.

2.	Firm-years are selected based on available CRSP stock return data. While @ball_how_2008 restricts the sample to firms with at least 240 trading days per year, I follow course guidance and do not impose this restriction, allowing for a slightly larger sample.

3.	Compustat includes active and inactive firms; since the original study does not filter based on firm activity, all firms meeting sample selection criteria (four quarterly earnings announcements per year, available stock returns) are retained.

4.	Permanent identifiers (`PERMNO`, `PERMCO`, `GVKEY`) ensure accurate historical tracking [@gow_empirical_2025]. `PERMNO` is used instead of `PERMCO` as it uniquely identifies individual securities rather than companies, preventing aggregation issues from multiple share classes.

5.	The replication relies on Compustat and CRSP, linked via the CRSP/Compustat Merged (CCM) database. As Compustat updates daily, while CRSP and CCM update annually, timing mismatches may arise. Unlike the original study, this replication uses the latest WRDS version as of January 2025, meaning database updates may introduce firm coverage or value differences.

6. Since @ball_how_2008 analyze returns, I use `RET` variable in CRSP too to simplify the process. Bid and ask prices are ignored.

7. Same as @ball_how_2008, I group earnings announcements based on their actual announcement dates (`RDQ` from Compustat) rather than the fiscal periods they cover. Since Compustat provides the precise earnings announcement date and CRSP aligns stock returns with trading days (`date` in CRSP), this approach ensures consistency. So, no adjustments are made to realign quarters.

By following the steps outlined in @sec-replication_steps1 and adhering to the assumptions made, I aim to replicate the results from @ball_how_2008. A thorough step-by-step approach, with each step clearly documented, helps to understand and verify the outputs.

## Replication Steps {#sec-replication_steps1}
The analysis involves gathering and filtering stock return and earnings announcement data, linking CRSP stock data with Compustat firm-level accounting data fundamentals, and estimating the relationship between annual returns and earnings-announcement window returns. The replication process follows a structured workflow of data pull, preparation, and analysis, ensuring consistency with the original study.

**Step 1: Pulling the Data and Managing the Databases**  
Data is pulled from WRDS using Python to integrate daily stock returns from CRSP (`crsp_a_stock.dsf`) with quarterly earnings announcement data from Compustat  
(`comp_na_daily_all.fundq`), linked via the Compustat-CRSP link table  
(`crsp_a_ccm.ccmxpf_lnkhist`). `permno` unique identifier matches stock returns to firm fundamentals via `gvkey` for exact firm/year match. Key variables include `ret` (daily stock return) and `rdq` (earnings announcement date), enabling event study analysis from 1972 to 2023. While WRDS provides extensive financial datasets, documentation on precise variable selection and dataset integration is limited, requiring additional validation and cross-referencing with prior research. The data is pulled for the entire range 1972-2023 required for Task 1 and 2 to prevent double preparation step.

**Step 2: Data Preparation**  
Although not finished, the plan would be as follows: The merged dataset would integrate stock return data from CRSP (`crsp_daily_stock_returns`), firm fundamentals from Compustat (`compustat_fundq_1972_2023`), and the CRSP-Compustat linking table (`linkdata_compustat_crsp.parquet`). The `PERMNO` identifier from CRSP and `GVKEY` from Compustat are connected via `LPERMNO` in the linking table, ensuring correct firm-year alignment. Unnecessary variables from linked table, such as `linkprim`, `liid`, `linktype`, and `usedflag`, are dropped. The resulting dataset would retain only daily stock returns (`ret`), earnings announcement dates (`rdq`), fiscal year-end indicators (`fyr`), and firm identifiers (`permno, gvkey`), used further for Buy-and-Hold Returns (BHR) calculation and regression analysis.

After defining the event windows for each earnings announcement day in Q1-Q4, retain only the firms that have four earnings announcements per year, such that each of them falls within the same calendar year. Additionally, for the subgroup replication of Table 2 Panels C-D classify the firms based on their fiscal year-end (`fyr`) to define whether firms have December fiscal year-end firm or in any another month. Control for the duplicate and missing firm/years.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
The analysis step computes the BHR returns and performs a cross-sectional regression for each distinct firm/year.

Since @ball_how_2008 do not explicitly describe their return computation, I compute Buy-and-Hold Returns (BHR) following the cumulative return formula as discussed in @gundersen_returns_2022. The BHR formula captures compounded returns over multiple days, reflecting the stock’s actual performance, which is used to replicate Table 1.

## Results {#sec-results1}
Since Tables and Figure replication not finished, this section presents only the original @fig-fig1_original.

# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
This Section extends the analysis to 2007–2023, a period defined by major economic events such as the 2008 financial crisis and the COVID-19 pandemic, which may have influenced the informativeness of earnings announcements. While the task is relatively straightforward in terms of programming—requiring adjustments to the sample period—it demands careful economic interpretation of observed trends, market shifts, and potential institutional details that may influence the relationship between earnings announcements and stock returns.

@@

## Results {#sec-results2}
The sample shift to 2007–2023 is expected to bring potential shifts in earnings announcement informativeness due to significant economic and regulatory events. The 2008 Global Financial Crisis likely disrupted earnings predictability, with heightened volatility and uncertainty increasing the market’s sensitivity to earnings reports. The COVID-19 pandemic 2020 introduced extreme market uncertainty, potentially amplifying the informativeness of earnings announcements as investors relied on them to gauge firm resilience and recovery. Additionally, the introduction of Regulation (EU) No 537/2014 may have influenced earnings informativeness by improving transparency, audit quality, and financial reporting compliance across EU countries. Differences in compliance and audit coverage could have led to variations in how earnings announcements reflect new information in stock prices, potentially contributing to cross-country disparities in earnings informativeness.

# Task 3 - Cross-Country Replication {#sec-task3}
Task 3 extends the analysis beyond the U.S. by replicating Figure 1 using 1972–2023 data from a non-U.S. market, examining how earnings informativeness varies across regulatory environments and market structures. This requires mapping CRSP/Compustat variables to their Worldscope/Datastream equivalents while adjusting for accounting standards, market liquidity, and institutional factors. Worldscope provides earnings announcement data, and Datastream covers stock returns. While the core methodology remains, careful data handling and interpretation are essential for cross-country analysis, particularly with Canada.

## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
Since this task employs a generalization and extension approach by applying the same methodology to a country different from original report, it requires adjustments for different databases and market structures. Based on the original paper, I do not gather analyst expectations or estimate earnings deviations for my replication. Instead, I only require stock return data and earnings announcement dates.

Canada was chosen for its quarterly reporting frequency, ensuring comparability with U.S. data, as most firms report four times per year, with some semi-annual reports [@short_short_2025]. Despite its smaller market size, Canada’s regulatory environment, liquidity, and institutional settings closely align with the U.S., making it an ideal candidate. Additionally, Canadian firms registered with the SEC can report under U.S. GAAP, making their financial disclosures fully compatible with U.S. firms.

In addition, I impose the following assumptions apart from those listed in @sec-research_design_assumptions1 to ensure generalization and extension approach due to usage of other databases or where @ball_how_2008 do not provide explicit guidance:

1. To ensure a smooth and conflict-free process, I handle the WSCP/Datastream workflow separately, preventing potential interference with the CRSP/Compustat pull process and improving debugging efficiency. Although a unified pull/prepare/do_analysis file for all databases would be ideal and streamline management, the large-scale nature of the extracted datasets demands modular approach.

2. Since I do not perform subgroup analysis for December and non-December fiscal year end firms as in Table 2 Panels C-D, `item5350` for FISCAL PERIOD END DATE is not relevant for this task, as Figure 1 applies to the analysis of the entire dataset. So, the sample is all firm-years with available data on the quarterly Worlscope and daily Datastream files.

3. I verified whether Worldscope and Datastream follow the same earnings announcement date convention as Compustat or if they report earnings based on fiscal periods. I decided to use Worldscope items 5901–5904 from historical stock-related data in `wrds_ws_stock`, which represent the actual date when a company publicly reports its earnings for each quarter, where, however, data for these fields is generally not available prior to 1992 [@thomson_financial_worldscope_2007, p.281]. This may cause limited data sample size.

4. This task is a generalization to a non-U.S. setting, meaning methodological differences will certainly arise due to variations in data structure and reporting standards. Worldscope - Stock Data `wrds_ws_stock` table is updated quarterly, while Datastream Daily Stock File `wrds_ds2dsf` is updated weekly, which may introduce timing mismatches when aligning financial fundamentals with stock returns. Given that, future applications should consider database updates and coverage differences between CRSP and Datastream.

5. Similar to @sec-task1, I use returns `ret` variable in Datastream for returns, presented as percentage change. Bid and ask prices are ignored.

6. I ignore the currency effect (CAD or USD) in DS table, as returns are measured in percentages, making them relative rather than absolute values.

@@

**Step 1: Pulling the Data and Managing the Databases**  
I follow the linking approach between Worldscope and Datastream based on @dai_research_2021. The original study uses data from 1972 onward, while Worldscope's `year_` variable only provides data starting from 1980 [@wharton_research_data_services_worldscope_2025]. This constraint may lead to differences in sample coverage. Hence, filter Datastream `marketdate` to 1980 to prevent mismatched observations where stock data exists but earnings do not.

Next, retrieve quarterly earnings announcement dates from Worldscope and daily stock returns from Datastream, filtering for Canadian firms from 1980 onward. The datasets are saved separately and later merged using firm-level identifiers. Only common equity (`typecode` = EQ) is included to ensure consistency with CRSP.

The pull step results into three comprehensive datasets: DS table with 27,950,012, Link table with 77,944, WSCP table with 1,942,202 observations.

**Step 2: Data Preparation**  
First, I merged Worldscope stock data with the linking table using `code` as unique identifier, resulting in 747,760 observations. This merged dataset was then joined with Datastream daily stock data using unique `infocode`, producing a final dataset with 10,249,737 observations.

If a firm had missing values for any of the Earnings Per Share Report Date fields (item5901–item5904), it did not meet the criteria set by @ball_how_2008 and was excluded from the sample. This filtering step removed 3,609,492 missing firm-year observations. Additionally, 5,996,793 firm-year entries were dropped because their four earnings announcements did not fall within the same calendar year. Furthermore, I extend dynamically shift the missing values for trading days -1,0,1 within 3 day window, due to small sample size to keep more earning announcemnet windows.

From this cleaned dataset, BHR event returns are computed. Then, annual stock returns for the firms of the filtered sample are extracted from Datastream pulled data . Finally, the Annual BHR Returns are calculated. Notably, after pulling the annual stock data for further annual return calculation, all infocodes present an average of 250 trading days, highlighting full year stock coverage.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
I conduct the earnings informativeness analysis by computing summary statistics (Table 1), running annual regressions (Table 2), and generating a replication of Figure 1. First, I loaded the computed event-window and annual buy-and-hold returns (BHR_3day, BHR_Annual) and produced summary statistics similar to @ball_how_2008. Next, I performed annual cross-sectional regressions of stock returns on earnings-announcement returns, benchmarking against an Abnormal R² (adjusted R² minus 4.8%), while handling missing data and ensuring a valid number of observations per year. Finally, the script plots Figure 1, visualizing Abnormal R² trends (Panel A) and slope coefficients (Panel B) over time, saving it as both a PNG and a pickle file for further use.

## Results {#sec-results3}
Comparing the replication results with those from @ball_how_2008 (see @fig-fig1_original) reveals both similarities and discrepancies. The original figure shows a general upward trend in abnormal  R² , peaking in the early 2000s, whereas the replicated figure captures greater volatility in the later years (2010–2023). Additionally, Ball’s figure suggests a smoother long-term dynamic, while the replication exhibits more pronounced fluctuations, particularly post-2010. Panel B in replicated figure exhibits higher volatility and larger coefficient magnitudes, suggesting increased earnings informativeness post-2006, while Ball’s original figure has more stable and lower-magnitude coefficients.

@@

The research design follows the generalization goal in empirical literature, emphasizing the need to validate findings across different methods [@bloomfield_gathering_2016]. @ball_how_2008 assess earnings informativeness through stock price reactions, but triangulation—beyond replication—requires diverse methods to examine the same question [@bloomfield_gathering_2016, p.353]. While price reactions indicate market response, they do not capture investor perceptions. A survey directly evaluates whether market participants view earnings announcements as confirmatory or informative. If investors confirm a secondary role, it reinforces Ball and Shivakumar’s (2008) findings, ensuring informativeness reflects cognitive and behavioral mechanisms rather than price fluctuations. These insights could inform corporate reporting strategies by clarifying how investors prioritize disclosures.

A survey is the most suitable non-archival method, capturing investor sentiment beyond what archival data provides. Unlike lab experiments, which lack real-world applicability, surveys measure the dependent variable (DV) by eliciting participant perceptions [@bloomfield_gathering_2016, p.358]. Field experiments, though capturing real-time reactions, require direct intervention, posing ethical and logistical challenges. A long-term research program combining surveys and field studies would enhance validity, leveraging unlimited funds to assess investor decision-making and market responses.

To ensure survey reliability, design strategies follow @bruggemann_worku_2024. A sample of 300 investors, analysts, and financial professionals will provide institutional and retail perspectives while balancing statistical power and feasibility. The survey will be distributed via email through financial networks, investor associations, and professional platforms (e.g., LinkedIn, CFA societies) and in paper format at finance conferences. The Humboldt-Universität zu Berlin (HU) brand may be used, subject to approval. Respondents will be assured confidentiality and offered an €20 honorarium. A pilot study with 30 participants will refine question clarity, survey length, and response distribution, helping identify unclear or missing questions and assess completion time.

>>>>

---
title: |
  | How Much New Information Is There in Earnings?
  | Corporate Decision-Making and Quantitative Analysis
  | Summer 2025 - Individual Report\vspace{1cm}
author:
  - name: Mel Mzv
    email: m@student.hu-berlin.de
    affiliations:
      - Humboldt-Universität zu Berlin  
date: today
date-format: MMM D, YYYY [\vspace{0cm}] # Adjusted the vertical space after the date from 1cm
abstract: |
  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide infrastructure for open science-oriented empirical projects. Leveraging data from the CRSP and Compustat databases, alongside Worldscope and Datastream via WRDS, this repository showcases a reproducible workflow integrating Python scripts for data preparation, analysis, and visualization. Integrating multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively. The project replicates and extends Ball and Shivakumar (2008) to analyze the informativeness of quarterly earnings announcements and their contribution to share price movements, highlighting their critical role in investment decisions and impact on investors and analysts. Key tasks include replicating and comparing original results, extending the analysis to 2007–2023, and applying the methodology to a non-U.S. country. The project also documents research design choices, discusses variations between original and reproduced results, and provides insights into earnings informativeness across different timeframes and jurisdictions. Additionally, it sketches a research design for a non-archival study to evaluate the paper’s findings. This code base, adapted from TREAT, demonstrates how the template applies to this project and serves as a structured guide for reproducible empirical research in accounting.
  <!--
  | \vspace{6cm} commenting away to decrease distance
  -->

bibliography: references.bib
biblio-style: apsr
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: true
    lof: true  # Enables List of Figures
    lot: true  # Enables List of Tables
fig_caption: yes
fontsize: 11pt
ident: yes
always_allow_html: yes
number-sections: true 
header-includes:
  - \usepackage[nolists]{endfloat}    
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=24pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
  - \usepackage[margin=1in]{geometry}
---

@@

# List of Abbreviations
\setlength{\parindent}{0em}
**BHR**: Buy-and-Hold Returns   
**CRSP**: Center for Research in Security Prices  
**DS**: Datastream  
**IDE**: Integrated Development Environment  
**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
**WRDS**: Wharton Research Data Services  
**WSCP**: Worldscope

@@

# Introduction {#sec-introduction}
This project serves as a template-based example for conducting and documenting reproducible empirical accounting research. It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis course, as well as the Accounting Reading Group course, which explored archival and non-archival research methods. To support future empirical thesis projects, this template was adapted and further developed specifically for students at the School of Business and Economics at Humboldt-Universität zu Berlin—particularly those affiliated with the Institute of Accounting and Auditing and the Finance Group.

Following prior assignments on audit market concentration and financial reporting effects (e.g., @graham_economic_2005), this project focuses on the informativeness of quarterly earnings announcements. Specifically, it replicates and extends the study by @ball_how_2008, which examines whether earnings announcements contribute new information to markets or merely confirm what is already known—contributing to the long-standing discussion on market efficiency and the valuation versus contracting roles of accounting.

Using updated U.S. data and applying the methodology to a second country, this thesis combines multiple WRDS datasets and applies an event-study design to evaluate how much of the annual stock return variation can be explained by earnings announcement windows. The approach is designed to foster transparent research workflows, with a fully scripted pipeline for data pulling, preparation, analysis, and reporting.

The paper is organized into four main parts. In @sec-task1, I replicate the core findings of the original study using U.S. data from 1972 to 2006. In @sec-task2, I apply the same methodology to U.S. data from 2007 to 2023 to explore temporal robustness. In @sec-task3, I generalize the findings to a non-U.S. country to assess cross-country variation and data integration requirements. Finally, @sec-task4 sketches a non-archival research design using a survey-based approach to evaluate the underlying assumptions of the original paper. The thesis concludes with reflections and key insights in @sec-conclusion.

# Task 1 - Replication of Key Tables and Figures {#sec-task1}
Task 1 replicates key tables and figures from @ball_how_2008 using daily US stock return data (1972–2006) to establish baseline results. By reconstructing Table 1 (Panel A), Table 2, and Figure 1, it evaluates the robustness of the original findings and highlights any discrepancies. Daily CRSP data ensures accurate measurement of stock reactions around earnings announcements. Replication results are discussed in @sec-results1, with extended analysis in @sec-task2.

## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
In line with @ball_how_2008, this replication examines U.S. earnings announcements and stock returns from 1972 to 2006, following the original study’s sample design, event window (-1 to +1), and return computations. Unlike the original, I omit analyst expectations, focusing solely on returns and announcement dates for simplicity.

To address gaps in methodological detail, several assumptions are applied. Event windows follow CRSP trading calendar conventions, shifting dates when earnings fall on non-trading days [@center_for_research_in_security_prices_crsp_crsp_2024]. While @ball_how_2008 require a minimum of 240 trading days, I relax this constraint, as per course guidance, resulting in a broader sample. All firms in Compustat meeting the quarterly announcement and return data criteria are included, regardless of activity status. I use PERMNO to uniquely track securities, avoiding aggregation across multiple share classes [@gow_empirical_2025]. The CRSP/Compustat Merged (CCM) database is used to link identifiers, noting that differences from the original study may arise due to updated data (WRDS version: January 2025). As in @ball_how_2008, the analysis is return-based (RET in CRSP), excluding bid-ask pricing. Earnings are grouped by announcement date (RDQ), not fiscal period, to align directly with CRSP trading dates.

These assumptions, together with the procedural details in @sec-replication_steps1, guide the replication of @ball_how_2008 and ensure transparency in design and implementation.

## Replication Steps {#sec-replication_steps1}
This section outlines the structured replication process: data pulling, preparation, and analysis, aligned with @ball_how_2008.

**Step 1: Pulling the Data and Managing the Databases**  
Data from WRDS combines daily CRSP stock returns (dsf) and Compustat quarterly earnings announcements (fundq), linked via the CRSP-Compustat link table (ccmxpf_lnkhist). The `permno` and `gvkey` identifiers ensure accurate firm-year alignment. Key variables are `ret` (returns) and `rdq` (announcement dates), pulled for the full 1972–2023 period used in Tasks 1 and 2.

**Step 2: Data Preparation**  
The merged dataset includes CRSP, Compustat, and the link table. After removing irrelevant linking variables, the dataset is trimmed to key variables: `ret`, `rdq`, `fyr`, `permno`, and `gvkey`. Only firms with four earnings announcements per calendar year are retained. Firms are also classified by fiscal year-end (`fyr`) to match Table 2 Panels C–D. Duplicate or incomplete firm-years are excluded.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
BHRs are computed using the cumulative return formula per @gundersen_returns_2022. Cross-sectional regressions are then run for each firm-year to replicate Table 1, Table 2, and Figure 1 from the original paper.

## Results {#sec-results1}
This section presents the replication results.

# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
This section extends the analysis to 2007–2023, a period shaped by events like the 2008 financial crisis and the COVID-19 pandemic. Although the programming involves only a sample period update, the task requires thoughtful economic interpretation of how structural shifts and evolving market conditions may have affected the informativeness of earnings announcements.

@@

## Results {#sec-results2}
Shifting the sample to 2007–2023 introduces potential changes in earnings announcement informativeness driven by key economic and regulatory events. The 2008 financial crisis and the 2020 COVID-19 pandemic likely heightened market uncertainty, increasing reliance on earnings disclosures to assess firm stability. Moreover, Regulation (EU) No 537/2014 may have enhanced financial reporting transparency and audit quality in EU countries, potentially affecting how earnings news is reflected in stock prices and contributing to cross-country variation in informativeness.

# Task 3 - Cross-Country Replication {#sec-task3}
Task 3 generalizes the analysis to a non-U.S. market by replicating Figure 1 using 1972–2023 data from Canada. This involves aligning CRSP/Compustat variables with Worldscope and Datastream equivalents while accounting for differences in reporting standards, market liquidity, and institutional context. Though the core methodology is retained, adapting it to international data demands careful mapping and interpretation.

## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
Since Task 3 applies a generalization and extension approach to a non-U.S. market, it requires adapting the methodology to alternative databases and market structures. Like in the original study, no analyst forecasts or earnings surprises are used—only stock returns and earnings announcement dates are required. Canada was selected for its comparable quarterly reporting frequency and institutional similarity to the U.S., making it a suitable choice for cross-country replication [@short_short_2025]. Canadian firms reporting under U.S. GAAP further enhance comparability.

To address database-specific issues not covered in @sec-research_design_assumptions1, I make the following adjustments: First, Worldscope/Datastream workflows are kept separate from CRSP/Compustat to prevent interference and facilitate debugging, given the scale of data involved. Second, I exclude fiscal-year-end-based subgroup analysis (Table 2 Panels C–D), as Figure 1 uses the full sample without splitting by fiscal period. Third, I use Worldscope items 5901–5904 to define earnings dates, though limited availability before 1992 may reduce the sample size [@thomson_financial_worldscope_2007, p.281]. Fourth, differences in update frequency between quarterly Worldscope and weekly Datastream can lead to timing mismatches when aligning fundamentals and returns. Fifth, I rely on Datastream’s `ret` variable for percentage returns, ignoring bid/ask spreads. Finally, currency effects are ignored, as percentage returns are unaffected by currency denomination.

@@

**Step 1: Pulling the Data and Managing the Databases**  
Following @dai_research_2021, I link Worldscope and Datastream using code and infocode. Since Worldscope’s year_ starts in 1980 [@wharton_research_data_services_worldscope_2025], Datastream data is also restricted to 1980–2023 to ensure consistent coverage. I extract quarterly earnings announcement dates (items 5901–5904) from Worldscope and daily stock returns from Datastream, focusing on Canadian firms and filtering for common equity (typecode = EQ). This step yields three datasets: 27.9M (Datastream), 77.9K (link table), and 1.94M (Worldscope) observations.

**Step 2: Data Preparation**  
Worldscope is first merged with the link table via code, producing 747,760 rows. This is then joined with Datastream via infocode, resulting in 10.2M observations. I exclude firms missing any earnings date fields, removing 3.6M rows, and drop 6M rows where earnings do not span all four quarters within the same calendar year. To retain more valid windows, event days (-1, 0, 1) are dynamically shifted within ±3 days when needed. After cleaning, I compute 3-day BHRs and extract full-year stock return data from Datastream. All retained infocode entries have around 250 trading days per year, confirming data completeness for annual BHR computation.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
The analysis replicates Table 1 (summary statistics), Table 2 (yearly regressions), and Figure 1 (Abnormal R² and slope trends). Using `BHR_3day` and `BHR_Annual`, I run yearly regressions of annual returns on earnings-window returns and compute Abnormal R² (adjusted R² – 4.8%), handling missing values appropriately. The resulting trends are plotted and saved as both PNG and pickle files for use in the paper.

## Results {#sec-results3}
Comparing the replication results with those from @ball_how_2008 reveals both similarities and discrepancies. The original figure shows a general upward trend in abnormal  R² , peaking in the early 2000s, whereas the replicated figure captures greater volatility in the later years (2010–2023). Additionally, Ball’s figure suggests a smoother long-term dynamic, while the replication exhibits more pronounced fluctuations, particularly post-2010. Panel B in replicated figure exhibits higher volatility and larger coefficient magnitudes, suggesting increased earnings informativeness post-2006, while Ball’s original figure has more stable and lower-magnitude coefficients.

@@

To extend the analysis beyond archival methods, this section proposes a survey to assess whether investors perceive earnings announcements as informative or merely confirmatory, following the triangulation principle in empirical research [@bloomfield_gathering_2016]. While @ball_how_2008 measure market reactions via stock returns, surveys can capture investor beliefs directly, offering complementary evidence on earnings informativeness.

A structured survey targeting 300 institutional and retail investors will be distributed through financial networks and platforms (e.g., LinkedIn, CFA societies), with a pilot study refining its clarity. Participants will receive a €20 honorarium, and responses will remain anonymous. The survey measures perceived informativeness (DV) using Likert-scale and multiple-choice questions, with IVs capturing factors such as investor preparedness and reaction timing.

Follow-up interviews with selected respondents will add qualitative depth, enabling a richer understanding of investor perspectives. This mixed-method approach enhances validity and supports generalization of findings beyond price-based evidence.
